---
title: "Week 2, Day 1 Exercises"
author: "Rob J Hyndman and George Athanasopoulos (Spencer Fox edits)"
---

```{r}
#| label: setup
#| include: false

## Caches files for quick loading
knitr::opts_chunk$set(
  cache = TRUE,
  cache.path = file.path("_cache", knitr::current_input(), "")
)

library(fpp3)
```

# fpp3 2.10, Ex 1

> Explore the following four time series: `Bricks` from `aus_production`, `Lynx` from `pelt`, `Close` from `gafa_stock`, `Demand` from `vic_elec`.
>
>   + Use `?` (or `help()`) to find out about the data in each series.
>   + What is the time interval of each series?
>   + Use `autoplot()` to produce a time plot of each series.
>   + Describe the trends, seasonal, and cyclical patterns for each
>   + For the `vic_elec`, modify the axis labels and title to be more accurate.

## Bricks

```{r}


```


## Lynx

```{r}


```


## Close

```{r}

```

## Demand

```{r}

```


# fpp3 2.10, Ex 2

> Use `filter()` to find what days corresponded to the peak closing price for each of the four stocks in `gafa_stock`.

```{r}

```

# fpp3 2.10, Ex 3

> Download the file `tute1.csv` from [the book website](http://OTexts.com/fpp3/extrafiles/tute1.csv), open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation. Read in the file, turn it into a tsibble object, and plot it to describe the trends.

```{r}

```

# fpp3 2.10, Ex 4

> The `USgas` package contains data on the demand for natural gas in the US.

>    a. Install the `USgas` package.
>    b. Create a tsibble from `us_total` with year as the index and state as the key.
>    c. Plot the annual natural gas consumption by state for the New England area (comprising the states of Maine, Vermont, New Hampshire, Massachusetts, Connecticut and Rhode Island).

```{r}


```


# fpp3 2.10, Ex 5

> a. Download `tourism.xlsx` from [the book website](http://OTexts.com/fpp3/extrafiles/tourism.xlsx) and read it into R using `read_excel()` from the `readxl` package.
> b. Create a tsibble which is identical to the `tourism` tsibble from the `tsibble` package.
> c. Find what combination of `Region` and `Purpose` had the maximum number of overnight trips on average.
> d. Create a new tsibble which combines the Purposes and Regions, and just has total trips by State.

```{r}


```
