---
title: "Week 3, Day 2 Solutions (Chapter 3)"
author: "Rob J Hyndman and George Athanasopoulos (Spencer Fox edits)"
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE)
library(fpp3)
library(here)
here::i_am("R/w3d2-solutions.qmd")

# Install seasonal package if you have not already done so
#install.packages("seasonal", repos = "https://cran.ms.unimelb.edu.au/")
library(seasonal)
```

# fpp3 3.7, Ex 1

> Consider the GDP information in `global_economy`. Plot the GDP per capita for each country over time. Which country has the highest GDP per capita? How has this changed over time?

```{r}
#| label: gdp_per_capita

global_economy |>
  autoplot(GDP / Population, alpha = 0.3) +
  guides(colour = "none") # turn off the legend else hard to see as there are many series
```

```{r}
avg_gdp_pc <- global_economy |>
  as_tibble() |>
  group_by(Country) |>
  summarise(
    # Average GDP per capita for each country
    gdp_pc = mean(GDP / Population, na.rm = TRUE),
    # Most recent GDP per capita for each country
    last = last((GDP / Population)[!is.na(GDP / Population)])
  )
top_n(avg_gdp_pc, 5, gdp_pc)
```

```{r}
max_gdp_pc <- global_economy |>
  semi_join(
    avg_gdp_pc |>
      filter(gdp_pc == max(gdp_pc, na.rm = TRUE)),
    by = "Country"
  )
```

```{r}
# install.packages("ggrepel")
# Using geom_label_repel() gives nicer label positions than geom_label()
# If the ggrepel package is not available, you can use geom_label() instead
library(ggrepel)
global_economy |>
  ggplot(aes(x = Year, y = GDP / Population, group = Country)) +
  geom_line(alpha = 0.3) +
  geom_line(colour = "red", data = max_gdp_pc) +
  geom_label_repel(
    aes(label = Country, x = 2020, y = last),
    data = top_n(avg_gdp_pc, 5, last),
  )
```

```{r}
global_economy |>
  mutate(gdp_pc = GDP / Population) |>
  as_tibble() |>
  group_by(Year) |>
  mutate(max_gdp_pc = max(gdp_pc, na.rm = TRUE)) |>
  ungroup() |>
  filter(
    gdp_pc == max_gdp_pc,
    Year >= 1970
  ) |>
  arrange(Year)
```

- Monaco currently has the highest GDP per capita, taking over from Liechtenstein.
- These two countries have had the highest GDP per capita for most of the last 50 years.
- The only years since 1970 where either Monaco or Liechtenstein did not have the highest GDP per capita were 1976 and 1977, when the United Arab Emirates just beat Monaco.

# fpp3 3.7, Ex 2

> For each of the following series, make a graph of the data. If transforming seems appropriate, do so and describe the effect.

> * United States GDP from `global_economy`
> * Slaughter of Victorian "Bulls, bullocks and steers" in `aus_livestock`
> * Victorian Electricity Demand from `vic_elec`.
> * Gas production from `aus_production`

## United States GDP

```{r}
us_economy <- global_economy |>
  filter(Country == "United States")
us_economy |>
  autoplot(GDP)
```

-   Trend appears exponential, a transformation would be useful.

```{r}
us_economy |>
  autoplot(box_cox(GDP, 0))
```

-   A log transformation (Box-Cox with $\lambda = 0$) appears slightly
    too strong.

```{r}
us_economy |>
  autoplot(box_cox(GDP, 0.3))
```

-   Using $\lambda = 0.3$ looks pretty good, the trend is now almost
    linear.

Let's see what guerrero's method suggests.

```{r}
us_economy |>
  features(GDP, features = guerrero)
```

Pretty close to $\lambda = 0.3$, let's see how it looks:

```{r}
us_economy |>
  autoplot(box_cox(GDP, 0.2819714))
```

-   More or less the same. Box-Cox transformations are usually
    insensitive to the choice of $\lambda$.

## Slaughter of Victorian "Bulls, bullocks and steers"

```{r}
vic_bulls <- aus_livestock |>
  filter(State == "Victoria", Animal == "Bulls, bullocks and steers")
vic_bulls |>
  autoplot(Count)
```

-   Variation in the series appears to vary slightly with the number of
    bulls slaughtered in Victoria.
-   A transformation may be useful.

```{r}
vic_bulls |>
  autoplot(log(Count))
```

-   A log transformation (Box-Cox $\lambda = 0$) appears to normalise
    most of the variation. Let's check with guerrero's method.

```{r}
vic_bulls |>
  features(Count, features = guerrero)
```

-   Pretty close, guerrero suggests $\lambda = -0.045$. This is close
    enough to zero, so it is probably best to just use a log
    transformation (allowing better interpretations).

## Victorian Electricity Demand

```{r}
vic_elec |>
  autoplot(Demand)
```

-   Seasonal patterns for *time of day* hidden due to density of ink.

-   *Day-of-week* seasonality just visible.

-   *Time-of-year* seasonality is clear with increasing variance in
    winter and high skewness in summer.

```{r}
vic_elec |>
  autoplot(box_cox(Demand, 0))
```

-   A log transformation makes the variance more even and reduces the
    skewness.

-   Guerrero's method doesn't work here as there are several types of
    seasonality.

## Australian Gas production

```{r}
aus_production |>
  autoplot(Gas)
```

-   Variation in seasonal pattern grows proportionally to the amount of
    gas produced in Australia. A transformation should work well here.

```{r}
aus_production |>
  autoplot(box_cox(Gas, 0))
```

-   A log transformation appears slightly too strong, where the
    variation in periods with smaller gas production is now larger than
    the variation during greater gas production.

```{r}
aus_production |>
  features(Gas, features = guerrero)
```

-   Guerrero's method agrees by selecting a slightly weaker
    transformation. Let's see how it looks.

```{r}
aus_production |>
  autoplot(box_cox(Gas, 0.1095))
```

Looking good! The variation is now constant across the series.

# fpp3 3.7, Ex 3

> Why is a Box-Cox transformation unhelpful for the `canadian_gas` data?

```{r}
canadian_gas |>
  autoplot(Volume) +
  labs(
    x = "Year", y = "Gas production (billion cubic meters)",
    title = "Monthly Canadian gas production"
  )
```

-   Here the variation in the series is not proportional to the amount
    of gas production in Canada.

-   When small and large amounts of gas is being produced, we can
    observe small variation in the seasonal pattern.

-   However, between 1975 and 1990 the gas production is moderate, and
    the variation is large.

-   Power transformations (like the Box-Cox transformation) require the
    variability of the series to vary proportionately to the level of
    the series.

# fpp3 3.7, Ex 4

> What Box-Cox transformation would you select for your retail data (from Exercise 8 in Section 2.10)?

```{r}
set.seed(12345678)
myseries <- aus_retail |>
  filter(
    `Series ID` == sample(aus_retail$`Series ID`, 1),
    Month < yearmonth("2018 Jan")
  )
myseries |>
  autoplot(Turnover) +
  labs(
    y = "Turnover (million $AUD)", x = "Time (Years)",
    title = myseries$Industry[1],
    subtitle = myseries$State[1]
  )
```

The variation appears proportional to the level of the series, so a Box-Cox transformation may be useful.

```{r}
myseries |>
  autoplot(box_cox(Turnover, 0)) +
  labs(
    title = myseries$Industry[1],
    subtitle = myseries$State[1]
  )
```

A log transformation (Box-Cox $\lambda = 0$) appears about right here.

Let's check what the Guerrero method would suggest.

```{r}
myseries |>
  features(Turnover, features = guerrero)
```

This is close to zero, so it supports our choice above.

Note: your series may be different and require a different transformation than what I have used here.

# fpp3 3.7, Ex 5

> For the following series, find an appropriate Box-Cox transformation in order to stabilise the variance. Tobacco from `aus_production`, Economy class passengers between Melbourne and Sydney from `ansett`, and Pedestrian counts at Southern Cross Station from `pedestrian`.

## Australian tobacco production

```{r}
aus_production |>
  autoplot(Tobacco)
```

-   This variation in this series appears to be mostly constant across
    different levels of the series.
-   If any transformation is required, it would be a weak one. This can
    be seen if a strong transformation (such as log) is used.

```{r}
aus_production |>
  autoplot(log(Tobacco))
```

-   Guerrero's method suggests that $\lambda = 0.926$ is appropriate.
    This is a very weak transformation, as it is close to 1 (probably
    best to not bother transforming this series).

```{r}
aus_production |>
  features(Tobacco, features = guerrero)
```

```{r}
aus_production |>
  autoplot(box_cox(Tobacco, 0.926))
```

-   This series appears very similar to the original. The transformation
    is having almost no effect.

## Economy passengers between Melbourne and Sydney

```{r}
ansett |>
  filter(Airports == "MEL-SYD", Class == "Economy") |>
  autoplot(Passengers) +
  labs(title = "Economy passengers", subtitle = "MEL-SYD")
```

-   The data does not appear to vary proportionally to the level of the
    series.

-   There are many periods in this time series (such as the strike and
    change in seat classes) that may need further attention, but this is
    probably better resolved with modelling rather than transformations.

## Pedestrian counts

```{r}
pedestrian |>
  filter(Sensor == "Southern Cross Station") |>
  autoplot(Count) +
  labs(title = "Southern Cross Pedestrians")
```

-   There is a high skewness and some zeros (so we can't take logs).
    Let's try the `log(x+1)` transformation:

```{r}
pedestrian |>
  filter(Sensor == "Southern Cross Station") |>
  autoplot(log1p(Count)) +
  labs(title = "Southern Cross Pedestrians")
```

That's roughly balanced the two tails.

# fpp3 3.7, Ex 6

> Show that a $3\times 5$ MA is equivalent to a 7-term weighted moving average with weights of 0.067, 0.133, 0.200, 0.200, 0.200, 0.133, and 0.067.

5-term moving average:
$$z_j = \frac{1}{5}(y_{j-2}+y_{j-1}+y_j+y_{j+1}+y_{j+2}).$$
3-term moving average:
$$u_t = \frac{1}{3}(z_{t-1}+z_t+z_{t+1}).$$
Substituting expression for $z_j$ into the latter formula we get
\begin{align*}
  u_t &= \frac{1}{3}\left(\frac{1}{5}\left(y_{t-3}+y_{t-2}+y_{t-1}+y_{t}+y_{t+1}\right)+\frac{1}{5}\left(y_{t-2}+y_{t-1}+y_t+y_{t+1}+y_{t+2}\right)+\frac{1}{5}\left(y_{t-1}+y_{t}+y_{t+1}+y_{t+2}+y_{t+3}\right)\right).\\
  &= \frac{1}{15}\left(y_{t-3}+2y_{t-2}+3y_{t-1}+3y_{t}+3y_{t+1}+2y_{t+2}+y_{t+3}\right),
\end{align*}
which is a 7-term weighted moving average with weights of 0.067, 0.133, 0.200, 0.200, 0.200, 0.133, and 0.067

# fpp3 3.7, Ex 7

> Consider the last five years of the Gas data from `aus_production`.

>    ```r
>    gas <- tail(aus_production, 5*4) |> select(Gas)
>    ```

>    a. Plot the time series. Can you identify seasonal fluctuations and/or a trend-cycle?

```{r}
gas <- tail(aus_production, 5 * 4) |> select(Gas)
gas |>
  autoplot(Gas) + labs(y = "Petajoules")
```

There is some strong seasonality and a trend.

>    b. Use `classical_decomposition` with `type=multiplicative` to calculate the trend-cycle and seasonal indices.
>    c. Do the results support the graphical interpretation from part a?

```{r}
decomp <- gas |>
  model(decomp = classical_decomposition(Gas, type = "multiplicative")) |>
  components()
decomp |> autoplot()
```

The decomposition has captured the seasonality and a slight trend.

>    d. Compute and plot the seasonally adjusted data.

```{r}
as_tsibble(decomp) |>
  autoplot(season_adjust) +
  labs(title = "Seasonally adjusted data", y = "Petajoules")
```

>    e. Change one observation to be an outlier (e.g., add 300 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?
>    f. Does it make any difference if the outlier is near the end rather than in the middle of the time series?

```{r}
gas |>
  mutate(Gas = if_else(Quarter == yearquarter("2007Q4"), Gas + 300, Gas)) |>
  model(decomp = classical_decomposition(Gas, type = "multiplicative")) |>
  components() |>
  as_tsibble() |>
  autoplot(season_adjust) +
  labs(title = "Seasonally adjusted data", y = "Petajoules")
```

-   The "seasonally adjusted" data now shows some seasonality because
    the outlier has affected the estimate of the seasonal component.

```{r}
gas |>
  mutate(Gas = if_else(Quarter == yearquarter("2010Q2"), Gas + 300, Gas)) |>
  model(decomp = classical_decomposition(Gas, type = "multiplicative")) |>
  components() |>
  as_tsibble() |>
  autoplot(season_adjust) +
  labs(title = "Seasonally adjusted data", y = "Petajoules")
```

The seasonally adjusted data now show no seasonality because the outlier is in the part of the data where the trend can't be estimated.

# fpp3 3.7, Ex 8

> Recall your retail time series data (from Exercise 8 in Section 2.10). Decompose the series using X11. Does it reveal any outliers, or unusual features that you had not noticed previously?

```{r}
set.seed(12345678)
myseries <- aus_retail |>
  filter(`Series ID` == sample(aus_retail$`Series ID`, 1))
decomp <- myseries |>
  model(x11 = X_13ARIMA_SEATS(Turnover ~ x11())) |>
  components()
decomp |> autoplot()
```

Two outliers are now evident in the "irregular" component --- in December 1995 and July 2010.

# fpp3 3.7, Ex 9

> Figures 3.16 and 3.17 show the result of decomposing the number of persons in the civilian labour force in Australia each month from February 1978 to August 1995.

>   a. Write about 3--5 sentences describing the results of the decomposition. Pay particular attention to the scales of the graphs in making your interpretation.

* The Australian labour force has been decomposed into 3 components (trend, seasonality, and remainder) using an STL decomposition.
* The trend element has been captured well by the decomposition, as it smoothly increases with a similar pattern to the data. The trend is of the same scale as the data (indicated by similarly sized grey bars), and contributes most to the decomposition (having the smallest scale bar).
* The seasonal component changes slowly throughout the series, with the second seasonal peak diminishing as time goes on -- this component is the smallest contribution original data (having the largest scale bar).
* The remainder is well-behaved until 1991/1992 when there is a sharp drop. There also appears to be a smaller drop in 1993/1994. There is sometimes some leakage of the trend into the remainder component when the trend window is too large. This appears to have happened here. It would be better if the recession of 1991-1992, and the smaller dip in 1993, were both included in the trend estimate rather than the remainder estimate. This would require a smaller trend window than what was used.
* In the bottom graph, the seasonal component is shown using a sub-series plot. December is the highest employment month, followed by March and September. The seasonal component changes mostly in March (with a decrease in the most recent years). July and August are the months with the next largest changes. The least changing is June with the rest are somewhere between these. December and September show increases in the most recent years.

>   b. Is the recession of 1991/1992 visible in the estimated components?

Yes. The remainder shows a substantial drop during 1991 and 1992 coinciding with the recession.

# fpp3 3.7, Ex 10

> This exercise uses the `canadian_gas` data (monthly Canadian gas production in billions of cubic metres, January 1960 -- February 2005).
>
>   a. Plot the data using `autoplot()`, `gg_subseries()` and `gg_season()` to look at the effect of the changing seasonality over time. What do you think is causing it to change so much?

```{r}
#| label: gas1

canadian_gas |> autoplot(Volume)
canadian_gas |> gg_subseries(Volume)
canadian_gas |> gg_season(Volume)
```

- The changes in seasonality are possibly due to changes in the regulation of gas prices.

>   b. Do an STL decomposition of the data. You will need to choose a seasonal window to allow for the changing shape of the seasonal component.

```{r}
#| label: gas2

fit <- canadian_gas |>
  model(STL(Volume)) |>
  components()

fit
names(fit)

fit |> autoplot()
```

>   c. How does the seasonal shape change over time? [Hint: Try plotting the seasonal component using `gg_season()`.]

```{r}
#| label: gas3

fit |> gg_season(season_year)
```

-   Here the changes are easier to see. Up to about 1990 there is strong
    seasonality with the greatest volume in the Canadian winter.
-   The seasonality increases in size over time. After 1990 the
    seasonality changes shape and appears to be driven partly by the
    month length near the end of the series.

>   d. Can you produce a plausible seasonally adjusted series?

```{r}
#| label: gas4

canadian_gas |>
  autoplot(Volume) +
  autolayer(fit, season_adjust, col = "blue")
```

>   e. Compare the results with those obtained using SEATS and X11. How are they different?

```{r}
#| label: gas5

# remember to load library(seasonal) before attempting this question!
canadian_gas |>
  model(X_13ARIMA_SEATS(Volume ~ seats())) |>
  components() |>
  autoplot()
canadian_gas |>
  model(X_13ARIMA_SEATS(Volume ~ x11())) |>
  components() |>
  autoplot()
```

-   Note that SEATS fits a multiplicative decomposition by default, so
    it is hard to directly compare the results with the other two
    methods.

-   The X11 seasonal component is quite similar to the STL seasonal
    component. Both SEATS and X11 have estimated a more wiggly trend
    line than STL.
