---
title: "Week 4, Day 1 Exercises (Chapter 4)"
author: "Rob J Hyndman and George Athanasopoulos (Spencer Fox edits)"
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE)
library(fpp3)
library(here)
here::i_am("R/w4d1.qmd")
```

# fpp3 4.6, Ex 1

> Write a function to compute the mean and standard deviation of a time series, and apply it to the `PBS` data. Plot the series with the highest mean, and the series with the lowest standard deviation.

```{r}

```

# fpp3 4.6, Ex 2

> Use `GGally::ggpairs()` to look at the relationships between the STL-based features for the `tourism` data. You might wish to change `seasonal_peak_year` and `seasonal_trough_year` to factors. Which is the peak quarter for holidays in each state?

```{r}

```

# fpp3 4.6, Ex 3

> Use a feature-based approach to look for outlying series in the `PBS` data. What is unusual about the series you identify as "outliers"?

```{r}


```

> **Reflection:**  
> What is unusual about the outlying series you identified?
