---
title: "Week 6, Day 1 Exercises"
author: "Spencer J. Fox"
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE)

library(tidyverse)
library(fable)
library(tsibble)
library(cowplot)

## Run these as needed
## install.packages("arrow", repos = c("https://apache.r-universe.dev", "https://cran.r-project.org"))
## install.packages("hubData", repos = c("https://hubverse-org.r-universe.dev", "https://cloud.r-project.org"))
## install.packages("hubUtils")
library(hubData)
library(hubUtils)
library(here)
here::i_am("R/w6d1-solutions.qmd")
```



# Connecting your R session to the hub and reading in the data
```{r}
## Make a relative path to where the hub location is on your machine
local_hub_path <- '../ATSF2026/'
## Gather the data
## Need to change the path to point to where you downloaded the repository
ts_con <- connect_target_timeseries(hub_path = local_hub_path)
ts_con |> collect() -> wili_ts


## Different ways to get the data if you prefer
## Downloading should always work as long as you have internet
wili_ts <- read_csv('https://raw.githubusercontent.com/sjfox/ATSF2026/refs/heads/main/target-data/time-series.csv')
## Change this path to the local path to your repository
wili_ts <- read_csv('../../ATSF2026/target-data/time-series.csv')

```

# ATSF Forecasting Ex 1

## Exercise 1: Convert to a tsibble and inspect the time series

Instructions: Identify the date, location, and value columns in `wili_ts`. Convert the data to a `tsibble`, then plot the time series for a small set of locations so you can see the overall behavior and any missingness.

```{r}
# Solution
# Inspect the column names to find the date, location, and value fields
glimpse(wili_ts)

# Convert to a tsibble with date index and location/target as keys
wili_tsibble <- wili_ts |>
  as_tsibble(index = target_end_date, key = c(location, target))

# Quick time-series plot to inspect overall trends and gaps
wili_tsibble |> autoplot()

```

## Exercise 2: Naive forecast for one location and a chosen forecast date

Instructions: Choose a single location, fit a naive model using all data up to a specific origin date, then forecast four weeks ahead. Plot the historical series and the forecast together.

```{r}
# List available locations so we can choose one
wili_tsibble |> 
  distinct(location)

# Pick one region and keep data up to a chosen origin date
# This origin date is in the validation period (Oct 2015-May 2017)
loc_tsibble <- wili_tsibble |> 
  filter(location == 'HHS Region 5',
         target_end_date < '2016-12-05')

# Fit a naive model and forecast 4 weeks ahead
loc_tsibble |>
  model(NAIVE(observation)) |>
  forecast(h = 4) -> loc_forecast

# Plot forecast plus historical data (wide view)
loc_forecast |> 
  autoplot(wili_tsibble |> 
             filter(target_end_date < '2017-01-15')) 
# Zoom in around the origin date for readability
loc_forecast |> 
  autoplot(wili_tsibble |> 
             filter(target_end_date > '2016-09-15',
               target_end_date < '2017-01-15')) 
```

## Exercise 3: Naive forecasts for multiple locations from the same date

Instructions: Pick a single origin date and fit naive models for all locations at once. Forecast forward and visualize them with small multiples.

```{r}
# Use all locations but keep data up to a shared origin date
# The origin date is in the validation period (Oct 2015-May 2017)
all_loc_tsibble <- wili_tsibble |> 
  filter(target_end_date < '2016-12-05')

# Fit a naive model for each location in parallel and forecast 4 weeks ahead
all_loc_tsibble |>
  model(NAIVE(observation)) |>
  forecast(h = 4) -> all_loc_forecast

# Plot the forecasts with small multiples by location
all_loc_forecast |> 
  autoplot(wili_tsibble |> 
             filter(target_end_date > '2016-09-15',
               target_end_date < '2017-01-15')) +
  facet_wrap(~location, nrow = 3)
```

## Exercise 4: Transform the data before forecasting

Instructions: Run a forecast on a transformed y-value such as a `log1p()`. Compare the transformed forecast with the original naive forecast.

```{r}
# Apply a variance-stabilizing transform before modeling
# Use the same origin date as Exercise 3
all_loc_tsibble |>
  model(NAIVE(log1p(observation))) |>
  forecast(h = 4) -> all_loc_forecast_tranformed

# Plot transformed forecasts (note: still on transformed scale)
all_loc_forecast_tranformed |> 
  autoplot(wili_tsibble |> 
             filter(target_end_date > '2016-09-15',
               target_end_date < '2017-01-15')) +
  facet_wrap(~location, nrow = 3)
```

## Exercise 5: Try a different model

Instructions: Fit a different model (for example, `SNAIVE` or `RW`) on the same multi-location time series and compare its forecast to the naive forecast.

```{r}
# Seasonal naive model using a yearly lag as the seasonal period
all_loc_tsibble |>
  model(SNAIVE(observation~lag('year'))) |>
  forecast(h = 4) -> all_loc_forecast_seasonal

# Plot seasonal naive forecasts by location
all_loc_forecast_seasonal |> 
  autoplot(wili_tsibble |> 
             filter(target_end_date > '2016-09-15',
               target_end_date < '2017-01-15')) +
  facet_wrap(~location, nrow = 3)

# Random walk with drift as an alternative benchmark
all_loc_tsibble |>
  model(RW(observation ~ drift())) |>
  forecast(h = 4) -> all_loc_forecast_rw

# Plot random-walk forecasts by location
all_loc_forecast_rw |> 
  autoplot(wili_tsibble |> 
             filter(target_end_date > '2016-09-15',
               target_end_date < '2017-01-15')) +
  facet_wrap(~location, nrow = 3)

```
