---
title: "Week 6, Day 1 Exercises"
author: "Spencer J. Fox"
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE)

library(tidyverse)
library(fable)
library(tsibble)
library(cowplot)

## Run these as needed
## install.packages("arrow", repos = c("https://apache.r-universe.dev", "https://cran.r-project.org"))
## install.packages("hubData", repos = c("https://hubverse-org.r-universe.dev", "https://cloud.r-project.org"))
## install.packages("hubUtils")
library(hubData)
library(hubUtils)
library(here)
here::i_am("R/w6d1.qmd")
```



# Connecting your R session to the hub and reading in the data
```{r}
## Make a relative path to where the hub location is on your machine
local_hub_path <- '../ATSF2026/'
## Gather the data
## Need to change the path to point to where you downloaded the repository
ts_con <- connect_target_timeseries(hub_path = local_hub_path)
ts_con |> collect() -> wili_ts

```

# ATSF Forecasting Ex 1

## Exercise 1: Convert to a tsibble and inspect the time series

Instructions: Identify the date, location, and value columns in `wili_ts`. Convert the data to a `tsibble`, then plot the time series for a small set of locations so you can see the overall behavior and any missingness.

```{r}

```

## Exercise 2: Naive forecast for one location and a chosen forecast date

Instructions: Choose a single location, fit a naive model using all data up to a specific origin date, then forecast four weeks ahead. Plot the historical series and the forecast together.

```{r}

```

## Exercise 3: Naive forecasts for multiple locations from the same date

Instructions: Pick a single origin date and fit naive models for all locations at once. Forecast forward and visualize them with small multiples.

```{r}

```

## Exercise 4: Transform the data before forecasting

Instructions: Run a forecast on a transformed y-value such as a `log1p()`. Compare the transformed forecast with the original naive forecast.

```{r}

```

## Exercise 5: Try a different model

Instructions: Fit a different model (for example, `SNAIVE` or `RW`) on the same multi-location time series and compare its forecast to the naive forecast.

```{r}

```
